\chapter{Acknowledgements}
\label{chap:ack}

This thesis is a result of almost two years 
of research undertaken at the University of Melbourne.
As my first major step into the world of academic
research, I can say that this project has been 
extremely challenging, whilst also tremendously rewarding.
During the final semester of the research project, my life
was filled with modelling, coding, debugging,
more debugging and writing.
Each of which was thoroughly enjoyable.

All the wonderful people which I was surrounded
by deserve a non-trivial share of the credit.
Their help and encouragement throughout the project has been invaluable,
and for that I am grateful.

First and foremost, I must thank the one who deserves it the most,
Alysson Costa, my supervisor.
Without his guidance, this project may have taken numerous wrong turns
along the way.
Alysson has previously studied both the problem I tackled
and the method I have used, and so his familiarity with the
areas provided a much needed level of understanding in the early days of the project.
I really appreciate the theoretical contributions that Alysson has
made, but possibly more importantly,
I truly appreciate his kindness.
He has an uncanny ability to find the humour in life
and seems to always have an extra smile to offer when one is needed.
It has been a pleasure to work with the most enthusiastic Operations
Researcher I've met.

My friends, both old and new, have filled these days 
of my life with jokes, pranks and joy; and I cannot thank them
enough for it.
I wish to thank Ai, Anupama, Ali, Ria, Lotte, Mark, Mim, 
Shian, John, Curtis, Heather, Scott, Lachlan, 
Georgia, Ben, Constance and Vincent.
In particular, Ria's help and friendship over the whole course
of my Masters needs a special mention.
Working together on the group projects during our course
has been nothing but a pleasure.
Her drive to become the very best she can --- at anything she puts her mind to ---
has been contagious.
Even in this thesis, her contributions to extending the
formulations of the CP sub-problems were pivotal in their successful performance.

I wish to express my gratitude to my sister, Margaret.
Miraculously, all the house work that I was unable to
assist with this semester disappeared without
request or mention: for this I am thankful.
She provided an immense amount of help by taking the time
out of her life to
assist me in the proof reading process.
I thank Margaret for ensuring the reader does not need
to endure a complete butchering of the English language.

The fellow members of my research group have also provided
wonderful feedback during the course of this project.
Without fail, they have been able to find new fruitful directions to take the project,
whilst also steering me away from the directions not filled with quite so much fruit.
I wish to thank Landir, Allen, Cheng, Gala, Ashwani, Simran, Pamela and Chathranee.

The entire experience has been terrific.

And I would like to disacknowledge my mortal enemy, Anupama Pilbrow.